HaploQA vs. R/qtl2 Haplotype Reconstructions

### 2022-12-23 draft_report.Rmd
updated documentations

The function automatically saves the computed haplotype reconstruction results to an RDS file, which you can find in the /results/RDS/ directory of this repo. Whenever the 'qtl2_file_gen' toggle is set to True to regenerate qtl2 input files for a cross, the according RDS files for that cross need to be manually removed from the /results/RDS/ directory for the haplotype reconstruction results to be properly updated, as the functions that calculate discordance and generate visualizations read the results from these RDS files.

@@

```{r, eval=F}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

  #gm_pct_res[[ind]] <- df
  gigamuga_results[[ind]] <- sample_gm_res
}
gigamuga_results <- readRDS(gm_res_dir)
saveRDS(gigamuga_results, gm_res_dir)
#write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)
```

@@

@@

As mentioned in the configuration file, the truth model for CC uses crosstype 'risib8', the truth model for DO uses crosstype 'do', the truth model for BXD F3 uses crosstype 'ail', and the truth model for F2 uses crosstype 'f2'. The executions of pipelines for the truth models are as shown below:

```{r}
do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = T, samples_gen = F)
f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

>>>>

The function automatically saves the computed haplotype reconstruction results to a number of RDS files, which you can find in the /results/RDS/ directory of this repo. Whenever the 'qtl2_file_gen' toggle is set to True to regenerate qtl2 input files for a cross, the according result RDS files for that cross need to be manually removed from the /results/RDS/ directory for the haplotype reconstruction results to be properly updated, as the functions that calculate discordance metrics and construct visualizations read the results from these RDS files.

@@

```{r, eval=F}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

@@

  #gm_pct_res[[ind]] <- df
  gigamuga_results[[ind]] <- sample_gm_res
}
gigamuga_results <- readRDS(gm_res_dir)
saveRDS(gigamuga_results, gm_res_dir)
#write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)

for (ind in names(gigamuga_results)) {
  print(ind)
  df <- ind_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
  gm_pct_res[[ind]] <- df
}
```

@@

To summarize the above results, for GigaMUGA, we selected 277 samples from DO, 218 samples from CC, 71 samples from F3, 34 samples from F2, and 288 mutant samples. For MiniMUGA, we selected 25 samples. Among which, DO and CC are complex crosses with 8 contributing founders, F3 and F2 are relatively simple two-founder crosses, whereas the number of contributing founders in the mutant crosses samples can vary from 2 to 8. This is to ensure there are sufficient evidences to evaluate the accuracy of haplotype reconstruction results for crosses of all computational sizes and complexity.

@@

As mentioned in the configuration file, the truth model for CC uses crosstype 'risib8', the truth model for DO uses crosstype 'do', the truth model for BXD F3 uses crosstype 'ail', and the truth model for F2 uses crosstype 'f2'. The executions of pipelines for the truth models can be initiated using the function, 'truth_model_reconstruction'. Same as the previous haplotype reconstruction functions, please set toggle 'samples_gen' = True if the genotyping data for individuals need to be (re)downloaded from HaploQA, and set to False otherwise. please set toggle 'qtl2_file_gen' = F if the input files for qtl2 read_cross need to be (re)generated, and False otherwise. Please also remove the existing result RDS files for the according cross (if any) whenever the qtl2 input files were regenerated to make sure the results are updated properly.

```{r, eval = F}
do_truth_results <- truth_model_reconstruction('DO', list_pheno, qtl2_file_gen = F, samples_gen = F)
cc_truth_results <- truth_model_reconstruction('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)
bxd_truth_results <- truth_model_reconstruction('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)
f2_truth_results <- truth_model_reconstruction('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```
